---
title: "Analysis"
author: "Juan Lamilla"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(sf)
library(countrycode)
library(dplyr)
library(ggplot2)
```

## Instructions

The task this week is to:

Read in global gender inequality data.

Join the global gender inequality index to spatial data of the World, creating a new column of difference in inequality between 2010 and 2019.

Share it with the World on GitHub.

Add you repository URL to the circulated spreadsheet.

Tip the countrycode R package will be helpful!

Tip the gender inequality has changed in the last year, you will find what you need in the “All composite indices and components time series (1990-2021)” dataset, the metadata file beneath it will explain what the columns are.

### Loading and Cleaning the Data

```{r loading}
# Load in global gender inequality as csv
global_gii_df <- read.csv('data/HDR21-22_Composite_indices_complete_time_series.csv')

# Remove all columns except country, iso3, gii_1990, gii_2021
global_gii_df <- global_gii_df[,c('iso3', 'country', 'gii_1990', 'gii_2021')]

# Remove all rows where gii_1990 or gii_2021 is NA
global_gii_df <- global_gii_df[!is.na(global_gii_df$gii_1990) & !is.na(global_gii_df$gii_2021),]

# Create a new column called gii_diff that is the difference between gii_1990 and gii_2021
global_gii_df$gii_diff <- global_gii_df$gii_2021 - global_gii_df$gii_1990

# Create a new column called gii_change that shows whether GII improved or worsened
global_gii_df$gii_change <- case_when(global_gii_df$gii_diff <= 0 ~ "Improved", TRUE ~ "Worsened", na.values = NA)

# Load in countries shp
country_shp <- st_read('data/World_Countries_Generalized.shp')

# Using countrycodes, add in iso3 column to countries shp based on the country column
country_shp$iso3 <- countrycode(country_shp$COUNTRY, 'country.name', 'iso3c')
```

### Joining the Data

```{r joining}
# Join global_gii_df gii_diff column to country_shp based on iso3 using left_join
country_shp <- left_join(country_shp, global_gii_df[,c('iso3', 'gii_2021', 'gii_change')], by = 'iso3')
```

### Mapping the Data

```{r mapping}

# Create a map of country_shp with GII
# Fill: gii_2021 using a orange gradient
# Color: gii_change using blue / red bicolor
ggplot(country_shp) +
  geom_sf(aes(fill = gii_2021, color = gii_change)) +
  
  # scale for fill
  scale_fill_gradient(
    low = "white", high = "orange", na.value = "gray90", name = "GII in 2021"
  ) +
  
  # scale for color
  scale_color_manual(
    values = c("skyblue", "red2"),
    na.translate = FALSE,
    name = "GII Change"
  ) + 
    
  # flip scale
  guides(fill = guide_colorbar(reverse = TRUE)) +
  
  theme_void() + 
  labs(title = "Global Gender Inequality Index difference between 1990 and 2021")

```

